```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}
# Randomly subsample CSV to 100000 records
# df <- read_csv("data/US_births(2018).csv")
# df.subsampled <- sample_n(df, 100000, replace = FALSE)
# write.csv(df.subsampled, "data/US_births(2018)_subsampled.csv", row.names = FALSE)
```


```{r}
df <- read_csv("data/US_births(2018)_subsampled.csv")
df
```

```{r}
model <- lm(DBWT ~ WTGAIN + SEX + PWgt_R + M_Ht_In + MAGER + BMI, data = df)
# model <- lm(DBWT ~ ., data = df)
summary(model)
```

# Introduction

Intuitively speaking, it is obvious that vehicle CO2 emissions are highly dependent on the vehicle itself. In this report, we explore the factors that affect a vehicle's CO2 emissions. We can find information about the model, transmission fuel type/consumption, and CO2 emissions of a diverse variety of vehicles. In this report, we leverage linear regression to find potential linear relationships. We will also assess the effectiveness of using a linear regression model to predict CO2 emissions for a particular vehicle. Lastly, we will evaluate how such a model can be used to decide how to lower the CO2 emission of existing vehicles.

# Data Description

The dataset used is compiled over 7 years, and is taken from the Canadian Government's official open data website.

1. Make: company of the vehicle [Categorical]
2. Car Model [Categorical]
3. Class of vehicle: describes the size of the vehicle depending on their utility, weight, and capacity [Categorical]
4. Engine size: size of engine used in Litre [Numerical]
5. Cylinders: number of cylinders in a vehicle [Numerical]
6. Transmission: transmission type with number of gears [Categorical]
7. Fuel type: type of fuel used [Categorical]
8. Fuel consumption: Fuel consumption in city roads (L/100 km) [Numerical]
9. Fuel consumption: Fuel consumption in highways (L/100 km) [Numerical]
10. Fuel consumption combined: The combined fuel consumption (55% city, 45% highway) is shown in L/100 km [Numerical]
11. Fuel consumption combined: The combined fuel consumption (55% city, 45% highway) is shown in mpg [Numerical]
12. CO2 emissions: shown in g/mk [Numerical]